Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-02 EconGrowth.Rnw
Cleaned up the code prior to beginning the process of adding energy terms to the Cobb-Douglas model.

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(){
  preds <- lapply(countryAbbrevs, cobbDouglasPredictions)
  out <- rbind(preds[["US"]], preds[["UK"]], preds[["JP"]], preds[["CN"]], preds[["ZA"]], preds[["SA"]], preds[["IR"]], preds[["TZ"]], preds[["ZM"]])
  colnames(out) <- c("predGDP")
  return(out)
}

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName){
  modelCD <- cobbDouglasModel(countryName)
  # Checks validity of the model. AIC stands for Akaike's Information Criterion.
  aicCD  <- AIC(modelCD, k=2)
  summaryCD <- summary(modelCD) # Gives the nls summary table.
  #print(summaryCD)
  ciCD <- confint(modelCD, level = ciLevel); ciCD # Calculates confidence intervals for the CD model.
  # Calculate beta and its confidence interval and report it.
  alpha <- as.numeric(coef(modelCD)["alpha"])
  beta <- 1.0 - alpha
  beta.est <- deltaMethod(modelCD, "1 - alpha"); beta.est # Estimates beta and its standard error (SE).
  # Now calculate a confidence interval on beta
  dofCD <- summaryCD$df[2]; dofCD # Gives the degrees of freedom for the model.
  tvalCD <- qt(ciHalfLevel, df = dofCD); tvalCD
  betaCICD <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estCD <- data.frame(lambda = coef(modelCD)["lambda"], alpha = coef(modelCD)["alpha"], beta = beta, gamma = 0)
  #print(estCD)
  row.names(estCD) <- "CD"
  #row.names(estCD) <- "Cobb-Douglas: $y = e^{\\lambda t}k^{\\alpha}l^{\\beta}$"
  # The [1] subscripts pick off the lower confidence interval
  lowerCD <- data.frame(lambda = ciCD["lambda","2.5%"], alpha = ciCD["alpha", "2.5%"], beta = betaCICD[1], gamma = NA) 
  row.names(lowerCD) <- "-95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCD <- data.frame(lambda = ciCD["lambda","97.5%"], alpha = ciCD["alpha", "97.5%"], beta = betaCICD[2], gamma = NA)
  row.names(upperCD) <- "+95% CI"
  # Now create the data for a table.
  dataCD <- rbind(upperCD, estCD, lowerCD)
  return(dataCD)
}

@@

<<CD_Params_Graph, fig.width=figWidth, fig.height=3.5, fig.cap="Cobb-Douglas (without energy) model parameters.", echo=FALSE>>=
createCDParamsGraph <- function(){
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))

>>>>

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(){
  out <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasPredictions)) #rbinds all the predictions together
  colnames(out) <- c("predGDP")
  return(out)
}

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName){
  modelCD <- cobbDouglasModel(countryName)
  # Checks validity of the model. AIC stands for Akaike's Information Criterion.
  aicCD  <- AIC(modelCD, k=2)
  summaryCD <- summary(modelCD) # Gives the nls summary table.
  ciCD <- confint(modelCD, level = ciLevel); ciCD # Calculates confidence intervals for the CD model.
  # Calculate beta and its confidence interval and report it.
  alpha <- as.numeric(coef(modelCD)["alpha"])
  beta <- 1.0 - alpha
  beta.est <- deltaMethod(modelCD, "1 - alpha"); beta.est # Estimates beta and its standard error (SE).
  # Now calculate a confidence interval on beta
  dofCD <- summaryCD$df[2]; dofCD # Gives the degrees of freedom for the model.
  tvalCD <- qt(ciHalfLevel, df = dofCD); tvalCD
  betaCICD <- with(beta.est, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estCD <- data.frame(lambda = coef(modelCD)["lambda"], alpha = coef(modelCD)["alpha"], beta = beta, gamma = 0)
  row.names(estCD) <- "CD"
  # The [1] subscripts pick off the lower confidence interval
  lowerCD <- data.frame(lambda = ciCD["lambda","2.5%"], alpha = ciCD["alpha", "2.5%"], beta = betaCICD[1], gamma = NA) 
  row.names(lowerCD) <- "-95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCD <- data.frame(lambda = ciCD["lambda","97.5%"], alpha = ciCD["alpha", "97.5%"], beta = betaCICD[2], gamma = NA)
  row.names(upperCD) <- "+95% CI"
  # Now create the data for a table.
  dataCD <- rbind(upperCD, estCD, lowerCD)
  return(dataCD)
}

@@

<<CD_Params_Graph, fig.width=figWidth, fig.height=3.5, fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.", echo=FALSE>>=
createCDParamsGraph <- function(){
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda, alpha, or beta), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))
